---
title: "R : Exploring Penalized Models for Handling High-Dimensional Survival Data"
author: "John Pauline Pineda"
date: "June 27, 2023"
output: 
  html_document:
    toc: true
    toc_depth: 3
    theme: readable
    highlight: tango
    css: doc.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **1. Table of Contents**
|
##  1.1 Sample Data
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.1, warning=FALSE, message=FALSE}

```

</details>

##  1.2 Data Quality Assessment
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.2, warning=FALSE, message=FALSE}

```

</details>

##  1.3 Data Preprocessing

###  1.3.1 Outlier
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.1, warning=FALSE, message=FALSE}

```

</details>

###  1.3.2 Zero and Near-Zero Variance
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.2, warning=FALSE, message=FALSE}

```

</details>

###  1.3.3 Collinearity
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.3, warning=FALSE, message=FALSE}

```

</details>

###  1.3.4 Linear Dependencies
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.4, warning=FALSE, message=FALSE}

```

</details>

###  1.3.5 Pre-Processed Dataset
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.5, warning=FALSE, message=FALSE}

```

</details>

## 1.4 Data Exploration
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.4, warning=FALSE, message=FALSE}

```

</details>

## 1.5 Penalized Survival Models

###  1.5.1 Least Absolute Shrinkage and Selection Operator (LASSO)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.1, warning=FALSE, message=FALSE}

```

</details>

###  1.5.2 Adaptive Least Absolute Shrinkage and Selection Operator (ALASSO)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.2, warning=FALSE, message=FALSE}

```

</details>

###  1.5.3 Elastic Net (ENET)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.3, warning=FALSE, message=FALSE}

```

</details>

###  1.5.4 Adaptive Elasticnet (AENET)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.4, warning=FALSE, message=FALSE}

```

</details>

###  1.5.5 Minimax Concave Penalty (MCP)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.5, warning=FALSE, message=FALSE}

```

</details>

###  1.5.6 Minimax Concave Penalty with L2 Regularization (MNET)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.6, warning=FALSE, message=FALSE}

```

</details>

###  1.5.7 Smoothly Clipped Absolute Deviation (SCAD)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.7, warning=FALSE, message=FALSE}

```

</details>

###  1.5.8 Smoothly Clipped Absolute Deviation with L2 Regularization (SNET)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.8, warning=FALSE, message=FALSE}

```

</details>

###  1.5.9 Fused Least Absolute Shrinkage and Selection Operator (FLASSO)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.9, warning=FALSE, message=FALSE}

```

</details>

##  1.6 Penalty Comparison Summary
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.6, warning=FALSE, message=FALSE}

```

</details>


# **2. References**

