---
title: "R : Exploring Penalized Models for Handling High-Dimensional Survival Data"
author: "John Pauline Pineda"
date: "June 27, 2023"
output: 
  html_document:
    toc: true
    toc_depth: 3
    theme: readable
    highlight: tango
    css: doc.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **1. Table of Contents**
|
##  1.1 Sample Data
|
| The <mark style="background-color: #EEEEEE;color: #FF0000">**peakVO2**</mark>  dataset from the  <mark style="background-color: #CCECFF">**randomForestSRC**</mark> package was used for this illustrated example.
|
| Preliminary dataset assessment:
|
| **[A]** 2231 rows (observations)
| 
| **[B]** 41 columns (variables)
|      **[B.1]** 1/41 survival event status = <span style="color: #FF0000">died</span> variable (factor)
|             **[B.1.1]** Event = <span style="color: #FF0000">died=1</span> 
|             **[B.1.2]** Censored or Non-Event = <span style="color: #FF0000">died=0</span>
|      **[B.2]** 1/41 survival time = <span style="color: #FF0000">ttodead</span> variable (numeric)
|      **[B.3]** 39/41 predictors = All remaining variables (26/39 factor + 13/39 numeric)

|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.1, warning=FALSE, message=FALSE}
##################################
# Loading R libraries
##################################
library(survival)
library(survminer)
library(mice)
library(foreign)
library(rms)
library(Hmisc)
library(VIM)
library(gridExtra)
library(finalfit)
library(knitr)
library(dplyr)
library(gtsummary)
library(tidyr)
library(purrr)
library(moments)
library(skimr)
library(caret)
library(corrplot)
library(randomForestSRC)
library(SurvMetrics)
library(pec)
library(hdnom)

##################################
# Loading source and
# formulating the train set
##################################
data(peakVO2)
peakVO2 <- as.data.frame(peakVO2)

##################################
# Converting survival time
# to discrete monthly data
##################################
peakVO2$ttodead <- ceiling(peakVO2$ttodead*12)
peakVO2.Source <- peakVO2

##################################
# Setting the appropriate data types
# for exploratory data analysis
##################################
peakVO2.Assessment <- (as.data.frame(sapply(peakVO2, function(x) max(x))))
names(peakVO2.Assessment) <- c("Max.Value")
peakVO2.Assessment$Variables <- rownames(peakVO2.Assessment)
peakVO2.Assessment

(peakVO2.FactorNames <- peakVO2.Assessment[peakVO2.Assessment[,1]==1,2])
peakVO2.Factor <- peakVO2[,peakVO2.FactorNames]
peakVO2.Factor <- as.data.frame(lapply(peakVO2.Factor, factor))

peakVO2.Numeric <- as.data.frame(peakVO2[,!names(peakVO2) %in% peakVO2.FactorNames])

peakVO2 <- data.frame(peakVO2.Numeric,peakVO2.Factor)

##################################
# Performing a general exploration of the data set
##################################
dim(peakVO2)
str(peakVO2)
summary(peakVO2)

##################################
# Formulating a data type assessment summary
##################################
PDA <- peakVO2
(PDA.Summary <- data.frame(
  Column.Index=c(1:length(names(PDA))),
  Column.Name= names(PDA), 
  Column.Type=sapply(PDA, function(x) class(x)), 
  row.names=NULL)
)

```

</details>

##  1.2 Data Quality Assessment
|
| Data quality assessment:
|
| **[A]** No missing observations noted for any variable.
|
| **[B]** Low variance observed for 10 variables with First.Second.Mode.Ratio>5.
|      **[B.1]** <span style="color: #FF0000">dilver</span> variable (factor)
|      **[B.2]** <span style="color: #FF0000">nifed</span> variable (factor)
|      **[B.3]** <span style="color: #FF0000">angioten.II</span> variable (factor)
|      **[B.4]** <span style="color: #FF0000">vasodilator</span> variable (factor)
|      **[B.5]** <span style="color: #FF0000">diuretic.loop</span> variable (factor)
|      **[B.6]** <span style="color: #FF0000">diuretic.thiazide</span> variable (factor)
|      **[B.7]** <span style="color: #FF0000">insulin</span> variable (factor)
|      **[B.8]** <span style="color: #FF0000">q.wave.mi</span> variable (factor)
|      **[B.9]** <span style="color: #FF0000">niddm</span> variable (factor)
|      **[B.10]** <span style="color: #FF0000">black</span> variable (factor)
|
| **[C]** No low variance observed for any variable with Unique.Count.Ratio<0.01.
|
| **[D]** No high skewness observed for any variable with Skewness>3 or Skewness<(-3).
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.2, warning=FALSE, message=FALSE}
##################################
# Loading dataset
##################################
DQA <- peakVO2

##################################
# Formulating an overall data quality assessment summary
##################################
(DQA.Summary <- data.frame(
  Column.Index=c(1:length(names(DQA))),
  Column.Name= names(DQA),
  Column.Type=sapply(DQA, function(x) class(x)),
  Row.Count=sapply(DQA, function(x) nrow(DQA)),
  NA.Count=sapply(DQA,function(x)sum(is.na(x))),
  Fill.Rate=sapply(DQA,function(x)format(round((sum(!is.na(x))/nrow(DQA)),3),nsmall=3)),
  row.names=NULL)
)

##################################
# Listing all descriptors
##################################
DQA.Descriptors <- DQA

##################################
# Listing all numeric Descriptors
##################################
DQA.Descriptors.Numeric <- DQA.Descriptors[,sapply(DQA.Descriptors, is.numeric)]

if (length(names(DQA.Descriptors.Numeric))>0) {
    print(paste0("There are ",
               (length(names(DQA.Descriptors.Numeric))),
               " numeric descriptor variable(s)."))
} else {
  print("There are no numeric descriptor variables.")
}

##################################
# Listing all factor Descriptors
##################################
DQA.Descriptors.Factor <- DQA.Descriptors[,sapply(DQA.Descriptors, is.factor)]

if (length(names(DQA.Descriptors.Factor))>0) {
    print(paste0("There are ",
               (length(names(DQA.Descriptors.Factor))),
               " factor descriptor variable(s)."))
} else {
  print("There are no factor descriptor variables.")
}

##################################
# Formulating a data quality assessment summary for factor Descriptors
##################################
if (length(names(DQA.Descriptors.Factor))>0) {

  ##################################
  # Formulating a function to determine the first mode
  ##################################
  FirstModes <- function(x) {
    ux <- unique(na.omit(x))
    tab <- tabulate(match(x, ux))
    ux[tab == max(tab)]
  }

  ##################################
  # Formulating a function to determine the second mode
  ##################################
  SecondModes <- function(x) {
    ux <- unique(na.omit(x))
    tab <- tabulate(match(x, ux))
    fm = ux[tab == max(tab)]
    sm = x[!(x %in% fm)]
    usm <- unique(sm)
    tabsm <- tabulate(match(sm, usm))
    ifelse(is.na(usm[tabsm == max(tabsm)])==TRUE,
           return("x"),
           return(usm[tabsm == max(tabsm)]))
  }

  (DQA.Descriptors.Factor.Summary <- data.frame(
  Column.Name= names(DQA.Descriptors.Factor),
  Column.Type=sapply(DQA.Descriptors.Factor, function(x) class(x)),
  Unique.Count=sapply(DQA.Descriptors.Factor, function(x) length(unique(x))),
  First.Mode.Value=sapply(DQA.Descriptors.Factor, function(x) as.character(FirstModes(x)[1])),
  Second.Mode.Value=sapply(DQA.Descriptors.Factor, function(x) as.character(SecondModes(x)[1])),
  First.Mode.Count=sapply(DQA.Descriptors.Factor, function(x) sum(na.omit(x) == FirstModes(x)[1])),
  Second.Mode.Count=sapply(DQA.Descriptors.Factor, function(x) sum(na.omit(x) == SecondModes(x)[1])),
  Unique.Count.Ratio=sapply(DQA.Descriptors.Factor, function(x) format(round((length(unique(x))/nrow(DQA.Descriptors.Factor)),3), nsmall=3)),
  First.Second.Mode.Ratio=sapply(DQA.Descriptors.Factor, function(x) format(round((sum(na.omit(x) == FirstModes(x)[1])/sum(na.omit(x) == SecondModes(x)[1])),3), nsmall=3)),
  row.names=NULL)
  )

}

##################################
# Formulating a data quality assessment summary for numeric Descriptors
##################################
if (length(names(DQA.Descriptors.Numeric))>0) {

  ##################################
  # Formulating a function to determine the first mode
  ##################################
  FirstModes <- function(x) {
    ux <- unique(na.omit(x))
    tab <- tabulate(match(x, ux))
    ux[tab == max(tab)]
  }

  ##################################
  # Formulating a function to determine the second mode
  ##################################
  SecondModes <- function(x) {
    ux <- unique(na.omit(x))
    tab <- tabulate(match(x, ux))
    fm = ux[tab == max(tab)]
    sm = na.omit(x)[!(na.omit(x) %in% fm)]
    usm <- unique(sm)
    tabsm <- tabulate(match(sm, usm))
    ifelse(is.na(usm[tabsm == max(tabsm)])==TRUE,
           return(0.00001),
           return(usm[tabsm == max(tabsm)]))
  }

  (DQA.Descriptors.Numeric.Summary <- data.frame(
  Column.Name= names(DQA.Descriptors.Numeric),
  Column.Type=sapply(DQA.Descriptors.Numeric, function(x) class(x)),
  Unique.Count=sapply(DQA.Descriptors.Numeric, function(x) length(unique(x))),
  Unique.Count.Ratio=sapply(DQA.Descriptors.Numeric, function(x) format(round((length(unique(x))/nrow(DQA.Descriptors.Numeric)),3), nsmall=3)),
  First.Mode.Value=sapply(DQA.Descriptors.Numeric, function(x) format(round((FirstModes(x)[1]),3),nsmall=3)),
  Second.Mode.Value=sapply(DQA.Descriptors.Numeric, function(x) format(round((SecondModes(x)[1]),3),nsmall=3)),
  First.Mode.Count=sapply(DQA.Descriptors.Numeric, function(x) sum(na.omit(x) == FirstModes(x)[1])),
  Second.Mode.Count=sapply(DQA.Descriptors.Numeric, function(x) sum(na.omit(x) == SecondModes(x)[1])),
  First.Second.Mode.Ratio=sapply(DQA.Descriptors.Numeric, function(x) format(round((sum(na.omit(x) == FirstModes(x)[1])/sum(na.omit(x) == SecondModes(x)[1])),3), nsmall=3)),
  Minimum=sapply(DQA.Descriptors.Numeric, function(x) format(round(min(x,na.rm = TRUE),3), nsmall=3)),
  Mean=sapply(DQA.Descriptors.Numeric, function(x) format(round(mean(x,na.rm = TRUE),3), nsmall=3)),
  Median=sapply(DQA.Descriptors.Numeric, function(x) format(round(median(x,na.rm = TRUE),3), nsmall=3)),
  Maximum=sapply(DQA.Descriptors.Numeric, function(x) format(round(max(x,na.rm = TRUE),3), nsmall=3)),
  Skewness=sapply(DQA.Descriptors.Numeric, function(x) format(round(skewness(x,na.rm = TRUE),3), nsmall=3)),
  Kurtosis=sapply(DQA.Descriptors.Numeric, function(x) format(round(kurtosis(x,na.rm = TRUE),3), nsmall=3)),
  Percentile25th=sapply(DQA.Descriptors.Numeric, function(x) format(round(quantile(x,probs=0.25,na.rm = TRUE),3), nsmall=3)),
  Percentile75th=sapply(DQA.Descriptors.Numeric, function(x) format(round(quantile(x,probs=0.75,na.rm = TRUE),3), nsmall=3)),
  row.names=NULL)
  )

}

##################################
# Identifying potential data quality issues
##################################

##################################
# Checking for missing observations
##################################
if ((nrow(DQA.Summary[DQA.Summary$NA.Count>0,]))>0){
  print(paste0("Missing observations noted for ",
               (nrow(DQA.Summary[DQA.Summary$NA.Count>0,])),
               " variable(s) with NA.Count>0 and Fill.Rate<1.0."))
  DQA.Summary[DQA.Summary$NA.Count>0,]
} else {
  print("No missing observations noted.")
}

##################################
# Checking for zero or near-zero variance Descriptors
##################################
if (length(names(DQA.Descriptors.Factor))==0) {
  print("No factor descriptors noted.")
} else if (nrow(DQA.Descriptors.Factor.Summary[as.numeric(as.character(DQA.Descriptors.Factor.Summary$First.Second.Mode.Ratio))>5,])>0){
  print(paste0("Low variance observed for ",
               (nrow(DQA.Descriptors.Factor.Summary[as.numeric(as.character(DQA.Descriptors.Factor.Summary$First.Second.Mode.Ratio))>5,])),
               " factor variable(s) with First.Second.Mode.Ratio>5."))
  DQA.Descriptors.Factor.Summary[as.numeric(as.character(DQA.Descriptors.Factor.Summary$First.Second.Mode.Ratio))>5,]
} else {
  print("No low variance factor descriptors due to high first-second mode ratio noted.")
}

if (length(names(DQA.Descriptors.Numeric))==0) {
  print("No numeric descriptors noted.")
} else if (nrow(DQA.Descriptors.Numeric.Summary[as.numeric(as.character(DQA.Descriptors.Numeric.Summary$First.Second.Mode.Ratio))>5,])>0){
  print(paste0("Low variance observed for ",
               (nrow(DQA.Descriptors.Numeric.Summary[as.numeric(as.character(DQA.Descriptors.Numeric.Summary$First.Second.Mode.Ratio))>5,])),
               " numeric variable(s) with First.Second.Mode.Ratio>5."))
  DQA.Descriptors.Numeric.Summary[as.numeric(as.character(DQA.Descriptors.Numeric.Summary$First.Second.Mode.Ratio))>5,]
} else {
  print("No low variance numeric descriptors due to high first-second mode ratio noted.")
}

if (length(names(DQA.Descriptors.Numeric))==0) {
  print("No numeric descriptors noted.")
} else if (nrow(DQA.Descriptors.Numeric.Summary[as.numeric(as.character(DQA.Descriptors.Numeric.Summary$Unique.Count.Ratio))<0.01,])>0){
  print(paste0("Low variance observed for ",
               (nrow(DQA.Descriptors.Numeric.Summary[as.numeric(as.character(DQA.Descriptors.Numeric.Summary$Unique.Count.Ratio))<0.01,])),
               " numeric variable(s) with Unique.Count.Ratio<0.01."))
  DQA.Descriptors.Numeric.Summary[as.numeric(as.character(DQA.Descriptors.Numeric.Summary$Unique.Count.Ratio))<0.01,]
} else {
  print("No low variance numeric descriptors due to low unique count ratio noted.")
}

##################################
# Checking for skewed Descriptors
##################################
if (length(names(DQA.Descriptors.Numeric))==0) {
  print("No numeric descriptors noted.")
} else if (nrow(DQA.Descriptors.Numeric.Summary[as.numeric(as.character(DQA.Descriptors.Numeric.Summary$Skewness))>3 |
                                               as.numeric(as.character(DQA.Descriptors.Numeric.Summary$Skewness))<(-3),])>0){
  print(paste0("High skewness observed for ",
  (nrow(DQA.Descriptors.Numeric.Summary[as.numeric(as.character(DQA.Descriptors.Numeric.Summary$Skewness))>3 |
                                               as.numeric(as.character(DQA.Descriptors.Numeric.Summary$Skewness))<(-3),])),
  " numeric variable(s) with Skewness>3 or Skewness<(-3)."))
  DQA.Descriptors.Numeric.Summary[as.numeric(as.character(DQA.Descriptors.Numeric.Summary$Skewness))>3 |
                                 as.numeric(as.character(DQA.Descriptors.Numeric.Summary$Skewness))<(-3),]
} else {
  print("No skewed numeric descriptors noted.")
}

```

</details>

##  1.3 Data Preprocessing

###  1.3.1 Outlier
|
| Outlier data assessment:
|
| **[A]** Outliers noted for 12 variables  with the numeric data visualized through a boxplot including observations classified as suspected outliers using the IQR criterion. The IQR criterion means that all observations above the (75th percentile + 1.5 x IQR) or below the (25th percentile - 1.5 x IQR) are suspected outliers, where IQR is the difference between the third quartile (75th percentile) and first quartile (25th percentile). Outlier treatment for numerical stability remains optional depending on potential model requirements for the subsequent steps.
|      **[A.1]** <span style="color: #FF0000">age</span> variable (33 outliers detected)
|      **[A.2]** <span style="color: #FF0000">resting.systolic.bp</span> variable (26 outliers detected)
|      **[A.3]** <span style="color: #FF0000">resting.hr</span> variable (9 outliers detected)
|      **[A.4]** <span style="color: #FF0000">bmi</span> variable (34 outliers detected)
|      **[A.5]** <span style="color: #FF0000">peak.rer</span> variable (102 outliers detected)
|      **[A.6]** <span style="color: #FF0000">peak.vo2</span> variable (40 outliers detected)
|      **[A.7]** <span style="color: #FF0000">interval</span> variable (25 outliers detected)
|      **[A.8]** <span style="color: #FF0000">bun</span> variable (116 outliers detected)
|      **[A.9]** <span style="color: #FF0000">sodium</span> variable (140 outliers detected)
|      **[A.10]** <span style="color: #FF0000">hgb</span> variable (83 outliers detected)
|      **[A.11]** <span style="color: #FF0000">glucose</span> variable (222 outliers detected)
|      **[A.12]** <span style="color: #FF0000">crcl</span> variable (82 outliers detected)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.1, warning=FALSE, message=FALSE}
##################################
# Loading dataset
##################################
DPA <- peakVO2

##################################
# Listing all predictors
##################################
DPA.Predictors <- DPA[,!names(DPA) %in% c("died","ttodead")]

##################################
# Listing all numeric predictors
##################################
DPA.Predictors.Numeric <- DPA.Predictors[,sapply(DPA.Predictors, is.numeric)]

##################################
# Identifying outliers for the numeric predictors
##################################
OutlierCountList <- c()

for (i in 1:ncol(DPA.Predictors.Numeric)) {
  Outliers <- boxplot.stats(DPA.Predictors.Numeric[,i])$out
  OutlierCount <- length(Outliers)
  OutlierCountList <- append(OutlierCountList,OutlierCount)
  OutlierIndices <- which(DPA.Predictors.Numeric[,i] %in% c(Outliers))
  boxplot(DPA.Predictors.Numeric[,i], 
          ylab = names(DPA.Predictors.Numeric)[i], 
          main = names(DPA.Predictors.Numeric)[i],
          horizontal=TRUE)
  mtext(paste0(OutlierCount, " Outlier(s) Detected"))
}

OutlierCountSummary <- as.data.frame(cbind(names(DPA.Predictors.Numeric),(OutlierCountList)))
names(OutlierCountSummary) <- c("NumericPredictors","OutlierCount")
OutlierCountSummary$OutlierCount <- as.numeric(as.character(OutlierCountSummary$OutlierCount))
NumericPredictorWithOutlierCount <- nrow(OutlierCountSummary[OutlierCountSummary$OutlierCount>0,])
print(paste0(NumericPredictorWithOutlierCount, " numeric variable(s) were noted with outlier(s)." ))

##################################
# Gathering descriptive statistics
##################################
(DPA_Skimmed <- skim(DPA.Predictors.Numeric))

###################################
# Verifying the data dimensions
###################################
dim(DPA.Predictors.Numeric)

```

</details>

###  1.3.2 Zero and Near-Zero Variance
|
| Zero and near-zero variance data assessment:
|
| **[A]** Low variance noted for 10 variables from the previous data quality assessment using a lower threshold.
|
| **[B]** Low variance noted for 2 variables using a preprocessing summary from the <mark style="background-color: #CCECFF">**caret**</mark> package. The <span style="color: #0000FF">nearZeroVar</span> method using both the <span style="color: #0000FF">freqCut</span> and <span style="color: #0000FF">uniqueCut</span> criteria set at 95/5 and 10, respectively, were applied on the dataset.
|      **[B.1]** <span style="color: #FF0000">dilver</span> variable (factor)
|      **[B.2]** <span style="color: #FF0000">nifed</span> variable (factor)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.2, warning=FALSE, message=FALSE}
##################################
# Loading dataset
##################################
DPA <- peakVO2

##################################
# Gathering descriptive statistics
##################################
(DPA_Skimmed <- skim(DPA))

##################################
# Identifying columns with low variance
###################################
DPA_LowVariance <- nearZeroVar(DPA,
                               freqCut = 95/5,
                               uniqueCut = 10,
                               saveMetrics= TRUE)
(DPA_LowVariance[DPA_LowVariance$nzv,])

if ((nrow(DPA_LowVariance[DPA_LowVariance$nzv,]))==0){
  
  print("No low variance predictors noted.")
  
} else {

  print(paste0("Low variance observed for ",
               (nrow(DPA_LowVariance[DPA_LowVariance$nzv,])),
               " numeric variable(s) with First.Second.Mode.Ratio>4 and Unique.Count.Ratio<0.10."))
  
  DPA_LowVarianceForRemoval <- (nrow(DPA_LowVariance[DPA_LowVariance$nzv,]))
  
  print(paste0("Low variance can be resolved by removing ",
               (nrow(DPA_LowVariance[DPA_LowVariance$nzv,])),
               " numeric variable(s)."))
  
  for (j in 1:DPA_LowVarianceForRemoval) {
  DPA_LowVarianceRemovedVariable <- rownames(DPA_LowVariance[DPA_LowVariance$nzv,])[j]
  print(paste0("Variable ",
               j,
               " for removal: ",
               DPA_LowVarianceRemovedVariable))
  }
  
  DPA %>%
  skim() %>%
  dplyr::filter(skim_variable %in% rownames(DPA_LowVariance[DPA_LowVariance$nzv,]))

  ##################################
  # Filtering out columns with low variance
  #################################
  DPA_ExcludedLowVariance <- DPA[,!names(DPA) %in% rownames(DPA_LowVariance[DPA_LowVariance$nzv,])]
  
  ##################################
  # Gathering descriptive statistics
  ##################################
  (DPA_ExcludedLowVariance_Skimmed <- skim(DPA_ExcludedLowVariance))
}

```

</details>

###  1.3.3 Collinearity
|
| High collinearity data assessment:
|
| **[A]** No high correlation > 95% noted for any variable pairs as confirmed using the preprocessing summaries from the <mark style="background-color: #CCECFF">**caret**</mark> and <mark style="background-color: #CCECFF">**lares**</mark> packages.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.3, warning=FALSE, message=FALSE}
##################################
# Loading dataset
##################################
DPA <- peakVO2

##################################
# Listing all predictors
##################################
DPA.Predictors <- DPA[,!names(DPA) %in% c("died","ttodead")]

##################################
# Listing all numeric predictors
##################################
DPA.Predictors.Numeric <- DPA.Predictors[,sapply(DPA.Predictors, is.numeric)]

##################################
# Visualizing pairwise correlation between predictors
##################################
DPA_CorrelationTest <- cor.mtest(DPA.Predictors.Numeric,
                       method = "pearson",
                       conf.level = .95)

corrplot(cor(DPA.Predictors.Numeric,
             method = "pearson",
             use="pairwise.complete.obs"), 
         method = "circle",
         type = "upper", 
         order = "original", 
         tl.col = "black", 
         tl.cex = 0.75,
         tl.srt = 90, 
         sig.level = 0.05, 
         p.mat = DPA_CorrelationTest$p,
         insig = "blank")



##################################
# Identifying the highly correlated variables
##################################
DPA_Correlation <-  cor(DPA.Predictors.Numeric, 
                        method = "pearson",
                        use="pairwise.complete.obs")
(DPA_HighlyCorrelatedCount <- sum(abs(DPA_Correlation[upper.tri(DPA_Correlation)]) > 0.95))

if (DPA_HighlyCorrelatedCount == 0) {
  print("No highly correlated predictors noted.")
} else {
  print(paste0("High correlation observed for ",
               (DPA_HighlyCorrelatedCount),
               " pairs of numeric variable(s) with Correlation.Coefficient>0.95."))
  
  (DPA_HighlyCorrelatedPairs <- corr_cross(DPA.Predictors.Numeric,
  max_pvalue = 0.05, 
  top = DPA_HighlyCorrelatedCount,
  rm.na = TRUE,
  grid = FALSE
))
  
}


if (DPA_HighlyCorrelatedCount > 0) {
  DPA_HighlyCorrelated <- findCorrelation(DPA_Correlation, cutoff = 0.95)
  
  (DPA_HighlyCorrelatedForRemoval <- length(DPA_HighlyCorrelated))
  
  print(paste0("High correlation can be resolved by removing ",
               (DPA_HighlyCorrelatedForRemoval),
               " numeric variable(s)."))
  
  for (j in 1:DPA_HighlyCorrelatedForRemoval) {
  DPA_HighlyCorrelatedRemovedVariable <- colnames(DPA.Predictors.Numeric)[DPA_HighlyCorrelated[j]]
  print(paste0("Variable ",
               j,
               " for removal: ",
               DPA_HighlyCorrelatedRemovedVariable))
  }
  
  ##################################
  # Filtering out columns with high correlation
  #################################
  DPA_ExcludedHighCorrelation <- DPA[,-DPA_HighlyCorrelated]
  
  ##################################
  # Gathering descriptive statistics
  ##################################
  (DPA_ExcludedHighCorrelation_Skimmed <- skim(DPA_ExcludedHighCorrelation))

}

```

</details>

###  1.3.4 Linear Dependencies
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.4, warning=FALSE, message=FALSE}

```

</details>

###  1.3.5 Pre-Processed Dataset
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.5, warning=FALSE, message=FALSE}

```

</details>

## 1.4 Data Exploration
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.4, warning=FALSE, message=FALSE}

```

</details>

## 1.5 Penalized Survival Models

###  1.5.1 Least Absolute Shrinkage and Selection Operator (LASSO)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.1, warning=FALSE, message=FALSE}

```

</details>

###  1.5.2 Adaptive Least Absolute Shrinkage and Selection Operator (ALASSO)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.2, warning=FALSE, message=FALSE}

```

</details>

###  1.5.3 Elastic Net (ENET)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.3, warning=FALSE, message=FALSE}

```

</details>

###  1.5.4 Adaptive Elasticnet (AENET)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.4, warning=FALSE, message=FALSE}

```

</details>

###  1.5.5 Minimax Concave Penalty (MCP)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.5, warning=FALSE, message=FALSE}

```

</details>

###  1.5.6 Minimax Concave Penalty with L2 Regularization (MNET)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.6, warning=FALSE, message=FALSE}

```

</details>

###  1.5.7 Smoothly Clipped Absolute Deviation (SCAD)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.7, warning=FALSE, message=FALSE}

```

</details>

###  1.5.8 Smoothly Clipped Absolute Deviation with L2 Regularization (SNET)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.8, warning=FALSE, message=FALSE}

```

</details>

###  1.5.9 Fused Least Absolute Shrinkage and Selection Operator (FLASSO)
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.9, warning=FALSE, message=FALSE}

```

</details>

##  1.6 Penalty Comparison Summary
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.6, warning=FALSE, message=FALSE}

```

</details>


# **2. References**

